---
title: "ElectionEconomy"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Regression is a method that allows researchers to summarize how predictions or average values of an outcome vary across individuals defined by a set of predictors. For example, Figure 1.1a shows the incumbent party’s vote share in a series of U.S. presidential elections, plotted vs. a measure of economic growth in the period leading up to each election year. 

Figure 1.1b shows a linear regression fit to these data. 

The model allows us to predict the vote—with some uncertainty—given the economy and under the assumption that future elections are in some way like the past.
Then we estimate the regression, y = a + bx + error:2 M1 <- stan_glm(vote ~ growth, data=hibbs)
And then we add the fitted line to our graph:
   abline(coef(M1), col="gray")
   
# Setup
```{r}
pacman::p_load(rstanarm, arm, rprojroot)
```

# ELECTIONS ECONOMY DATA
```{r}
getwd()

# Get data
hibbs <- read.table("/Users/pernillebrams/Desktop/UNI/4th semester/CompMod/ROS-Examples/ElectionsEconomy/data/hibbs.dat", header = TRUE)
View(hibbs)

# Scatterplot
plot(hibbs$growth, hibbs$vote, xlab = "Avg recent growth in personal income", ylab = "Incumbent party's vote share")

# Estimate regression y = a + bx + error: 
model <- stan_glm(vote ~ growth, data = hibbs)

# Add fitted line to graph (have to run the two below simultaneously)
plot(hibbs$growth, hibbs$vote, xlab = "Avg recent growth in personal income", ylab = "Incumbent party's vote share")
abline(coef(model),col = "red")

# Display fitted model
print(model)
```
First column shows the coefficients of the fitted line: y = ax + b in the sense that: y = 3x + 46.3
Second column shows the uncertainties in the estimates using median absolute deviations.
Last line of output shows estimate and uncertainty of sigma, σ, the scale of variation in the data that is UNEXPLAINED by the regression model.

# SIMPLE CAUSAL DATA
## Simulated data from linear model
```{r}
set.seed(1151)
N <- 50
x <- runif(N, 1, 5)
y <- rnorm(N, 10 + 3*x, 3)
x_binary <- ifelse(x<3, 0, 1)
data <- data.frame(N, x, y, x_binary)

```

## Regression with binary predictor
```{r}
lm_1a <- lm(y ~ x_binary, data = data)
display(lm_1a)
```

## Plots
```{r}
par(mar=c(3, 3, 1.5, 1), mgp=c(1.7, .5, 0), tck=-.01)
plot(x_binary, y, xlab="", ylab="Outcome measurement", pch=20, cex=.5, bty="l", main="Regression with binary treatment", cex.main=.9, xaxt="n", cex.lab=.9, cex.axis=.9)
axis(1, c(0,1), c("    Control", "Treatment    "), cex.axis=.9)
abline(coef(lm_1a)[1], coef(lm_1a)[2])
text(0.3, 13, paste("Estimated treatment effect is\nslope of fitted line: ", fround(coef(lm_1a)[2], 1)), cex=.8, adj=0)
```
## Regression with continuous predictor
```{r}
lm_1b <- lm(y ~ x, data = data)
display(lm_1b)
```
## Plots
```{r}
par(mar=c(3, 3, 1.5, 1), mgp=c(1.7, .5, 0), tck=-.01)
plot(x, y, xlab="Treatment level", ylab="Outcome measurement", pch=20, cex=.5, bty="l", main="Regression with continuous treatment", cex.main=.9, cex.lab=.9, cex.axis=.9)
abline(coef(lm_1b)[1], coef(lm_1b)[2])
text(3.2, 15, paste("Estimated treatment\neffect per unit of x is\nslope of fitted line: ", fround(coef(lm_1b)[2], 1)), cex=.8, adj=0)
```
## Classical regression with continuous predictor
```{r}
lm_2a <- lm(y ~ x, data = data)
display(lm_2a)
```
## Plots
```{r}
par(mar=c(3, 3, 1.5, 1), mgp=c(1.7, .5, 0), tck=-.01)
plot(x, y, xlab="Treatment level", ylab="Outcome measurement", pch=20, cex=.5, bty="l", main="Nonlinear treatment effect", cex.main=.9, cex.lab=.9, cex.axis=.9)
curve(5 + 30*exp(-x), add=TRUE)

par(mar=c(3, 3, 1.5, 1), mgp=c(1.7, .5, 0), tck=-.01)
plot(x, y, xlab="Treatment level", ylab="Outcome measurement", pch=20, cex=.5, bty="l", main="Nonlinear effect, estimated with straight line fit", cex.main=.9, cex.lab=.9, cex.axis=.9)
abline(coef(lm_2a)[1], coef(lm_2a)[2])
```

# Simulated data from two groups
```{r}
set.seed(1151)
N <- 100
xx <- rnorm(N, 0, 1)^2
z <- rep(0:1, N/2)
xx <- ifelse(z==0, rnorm(N, 0, 1.2)^2, rnorm(N, 0, .8)^2)
yy <- rnorm(N, 20 + 5*xx + 10*z, 3)
data <- data.frame(xx, z, yy)
lm_2 <- lm(yy ~ xx + z, data=data)
display(lm_2)

```

## Plots
```{r}
par(mar=c(3, 3, 1.5, 1), mgp=c(1.7, .5, 0), tck=-.01)
plot(xx, yy, xlab="Pre-treatment predictor", ylab="Outcome measurement", bty="l", main="Continuous pre-treatment predictor and binary treatment    ", cex.main=.9, cex.lab=.9, cex.axis=.9, type="n")
points(xx[z==0], yy[z==0], pch=20, cex=.5)
points(xx[z==1], yy[z==1], pch=1, cex=1)
abline(coef(lm_2)[1], coef(lm_2)[2])
abline(coef(lm_2)[1] + coef(lm_2)[3], coef(lm_2)[2])
text(2.3, 29.5, "Controls", cex=.9, adj=0)
text(1.5, 45, "Treated", cex=.9, adj=0)
x0 <- 5.2
arrows(x0, coef(lm_2)[1] + coef(lm_2)[2]*x0, x0, coef(lm_2)[1] + coef(lm_2)[2]*x0 + coef(lm_2)[3], length=.1, code=3)
text(x0+.15, 1 + coef(lm_2)[1] + coef(lm_2)[2]*x0 + .5*coef(lm_2)[3], paste("Estimated\ntreatment\neffect is", fround(coef(lm_2)[3], 1)), cex=.8, adj=0)
```

